Towards a Mechanized Proof of Noninterference

### 2013-12-18 paper/paper.tex
A bit of fleshing out of the operational semantics section and small tweaks elsewhere.

In this paper we discuss our experience trying to mechanize a proof of
noninterference for a simple language with security labels, \langName.  Although
we did not complete the full noninterference proof because we ran out of time,
we were able to find small inaccuracies and subtleties in the proofs from the
paper we based our proof on.  We discuss what would be involved in completing
the proof and how this relates to other work.

@@

\textbf{TODO: Talk about what the bracketed expressions and values are.}

@@

@@

\begin{figure}[h!]
  \begin{align}
    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \: |
    \: <v | v> \\
    a & ::= v \tag{answers} \: | \: <a | a> \\
    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e \: e \: | \: \textsf{let
    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e> \\
  \end{align}
  \caption{Syntax of \langName}
  \label{fig:syntax}
\end{figure}

@@

\langName has an operational semantics that should be familiar to the reader who
has seen a single step semantics for a basic lambda calculus.  The interesting
additions are the same as those in \origLang, where there are additional rules
for handling the bracketed values.

\textbf{TODO: Talk about what the rules do at a high level.}

@@

>>>>

In this paper we discuss our experience trying to mechanize a proof of
noninterference for a simple language with security labels, \langName.  This
marks the first step towards writing a verified compiler for a language with
information flow types.  Although we did not complete the full noninterference
proof because we ran out of time, we were able to find small inaccuracies and
subtleties in the proofs from the paper we based our proof on.  We discuss what
would be involved in completing the proof and how this relates to other work.

@@

@@

\langName, like \origLang, includes angle bracketed expressions, which represent
values or expressions which could differ between two runs of the program.  As we
will see when discussing the type system, this corresponds to values and
expressions that have high security labels on their types.

@@

\begin{figure}[h!]
  \begin{align}
    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \: |
    \: <v | v> \\
    a & ::= v \tag{answers} \: | \: <a | a> \\
    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e \: e \: | \: \textsf{let
    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e>
  \end{align}
  \caption{Syntax of \langName}
  \label{fig:syntax}
\end{figure}

@@

\langName has an operational semantics that should be familiar to the reader who
has seen a single step semantics for a basic lambda calculus.  The interesting
additions are the same as those in \origLang, where there are additional rules
for handling the bracketed values.  The reduction rules for the semantics are
given in Figure~\ref{fig:operational_semantics}.

@@

At a high level, these bracketed expressions are evaluated by evaluating one of
the branches, as shown in rules \ref{eq:bracket-left} and
\ref{eq:bracket-right}, or by expanding the brackets around the parent
expression, passing the appropriate branch of the bracketed subvalue to each
branch of the resulting expression, as shown in rules \ref{eq:lift-app} and
\ref{eq:lift-if}.  In these rules, we use the notation $\lift{x}_i$ to denote
the left ($i = 1$) or right ($i = 2$) branch of a bracketed expression or value.

We prove soundness and completeness lemmas for these rules in
Section~\ref{sec:proof}.
